---
title: "Download.nc.files.from.CDS.with.cdsDownload()"
author: "Ugo Chiavetta and Sebastian Marzini"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Download.nc.files.from.CDS.with.cdsDownload()}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## 1.Create an account in Copernicus Climate Data Store 

Before downloading the nc file, the registration to the Copernicus CDS site is required (https://cds.climate.copernicus.eu/#!/home). After that be sure to accept  terms and conditions here: https://cds.climate.copernicus.eu/cdsapp/#!/terms/licence-to-use-copernicus-products while you are logged in.

After logging in, click on your name in the top right corner, scroll the page and take note of 
'UID' and 'API Key'. Then run the code below

```{r variables_needed, eval=FALSE}
user_id <- "XXXXX" #change XXXXX with your UID 
user_key <- "XXXXXXXX-XXXX-XXXX-XXXX-XXXXXXXXXXXX" #change XXXXX with your API Key 
N <- 44.3 #try to put known latitude
E <- 11.2 #try to put known longitude
dir <- getwd()
cdsDownload(U_ID = user_id, API_Key = user_key, lat = N, lon = E, sPath = dir, site_id = "")

```


Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
